aa_mic_submission_analyses

### 2021-09-16 code/aa_mic_submission_analyses.Rmd
final fig tweaking

{} to match other plot, facet by utility condition, color by agent type?
```{r}
rl_comparison <- d.by_round_metrics %>%
  group_by(languageCondition, condition, gameId) %>%
  summarize(avgRoundPoints = mean(roundUtility)) %>% 
  mutate(rl_agent = "Human") %>%
  rbind(d.rl.game_metrics) %>%
  ggplot(aes(x = languageCondition, y = avgRoundPoints, color = condition)) + 
  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + 
  scale_color_bay()+ theme_jmank() + 
 # labs(y = "Average Points", color = "Player Type") +
  labs(y = "Average Points per Round", x = "Language Condition", color = "Utility Condition") +
  facet_wrap(~rl_agent)+ 
  theme(legend.position = "bottom")

@@

## Points x Time x Num. Words
```{r}
lang_by_points <- d.by_game_metrics %>% 
  ggplot(aes(x = gameLength/60, y = gameUtility, shape = languageCondition, color = log(numWords))) + 
  geom_point(alpha = .8) +
  geom_smooth(aes(group = condition), method=lm, formula=y~log(x),color = "black")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Game length (in Minutes)", 
       y = "Points Earned by Group",
       color = "Log N. Words",
       shape = "Language") + 
  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555") + theme(legend.position = "bottom")
lang_by_points
```

@@

max(d.by_game_metrics$numWords)
```
# Language Reduction
Want number of words exchanged over the game
 More utterances in individual, (same turns)

 shared utilities

 more questions in shared?
```{r}
lang_reduction <- d.by_round_metrics %>%
  filter(languageCondition == "Lang") %>%
  ggplot(aes(x = trialNum, y = numWords, color = condition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  #geom_smooth(method=lm, formula=y~log(x))+
   labs(x = "Trial Number", y = "N. Words", color = "Utility Condition")+ 
  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom")# + scale_y_continuous(trans='log10')

@@

expand_utt <- d.raw_chat.final %>%
  select(gameId, trialNum, playerId) %>% 
  distinct() %>%
  full_join(utterances)

@@

#points increase as language decreases?
d.by_round_metrics %>%
  filter(languageCondition == "Lang") %>%
  ggplot(aes(x = trialNum, y = roundUtility, color = condition)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method=lm)+
  scale_color_bay() + theme_jmank()

@@

```{r}
library(patchwork)
lang_by_points
rl_comparison
lang_reduction

>>>>

{} to match other plot, facet by utility condition, color by agent type?
```{r}
rl_comparison <- d.by_round_metrics %>%
  group_by(languageCondition, condition, gameId) %>%
  summarize(avgRoundPoints = mean(roundUtility)) %>% 
  mutate(rl_agent = "Human") %>%
  rbind(d.rl.game_metrics) %>%
  ggplot(aes(x = languageCondition, y = avgRoundPoints, color = condition)) + 
  geom_point(position=position_jitterdodge(dodge.width=.5, jitter.width = .1), alpha = .25) +
  stat_summary(fun.data = "mean_cl_boot", position=position_dodge(width=.5)) + 
  scale_color_bay()+ theme_jmank() + 
 # labs(y = "Average Points", color = "Player Type") +
  labs(y = "Average Points per Trial", x = "Language Condition", color = "Utility Condition") +
  facet_wrap(~rl_agent)+ 
  theme(legend.position = "bottom",
        #legend.box="vertical",
        legend.margin=margin(),
        text = element_text(size=15),
        axis.text = element_text(size = 12),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        strip.text = element_text(size=15))

@@

## Points x Time x Num. Words
```{r}
lang_by_points <- d.by_game_metrics %>% 
  ggplot(aes(x = gameLength/60, y = gameUtility, shape = languageCondition, color = log(numWords))) + 
  geom_point(alpha = .8, size = 4) +
  geom_smooth(aes(group = condition), method=lm, formula=y~log(x),color = "black")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Game length (in Minutes)", 
       y = "Points Earned by Group",
       color = "Log N. Words",
       shape = "") + 
  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555") + 
  theme(legend.position = "bottom",
        #legend.box="vertical",
        legend.margin=margin(),
        text = element_text(size=15),
        axis.text = element_text(size = 12),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        strip.text = element_text(size=15))
lang_by_points
```

@@

max(d.by_game_metrics$numWords)
```
# Language Reduction

```{r}
# number of words exchanged decreases over time
lang_reduction <- d.by_round_metrics %>%
  filter(languageCondition == "Lang") %>%
  mutate(utility_cond_short = ifelse(condition == "Shared Utilities", "Shared", "Individual")) %>%
  ggplot(aes(x = trialNum, y = numWords, color = utility_cond_short)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  #geom_smooth(method=lm, formula=y~log(x))+
   labs(x = "Trial Number", y = "N. Words", color = "Utility Condition")+ 
  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom") + 
   theme(legend.position = "bottom",
        text = element_text(size=15),
        axis.text = element_text(size = 12),
        legend.title=element_text(size=12),
        legend.text=element_text(size=12),
        strip.text = element_text(size=15))
lang_reduction

@@

# normal MLU also decreases
utt_mlu <- utterances %>% 
  unnest(c(utt_length_chars, utt_length_words)) %>%
  left_join(d.games.conditions) %>%
  group_by(trialNum, gameId) %>%
  mutate(mlu = mean(utt_length_words)) %>% 
  ggplot(aes(x = trialNum, y = mlu, color = condition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  labs(x = "Trial Number", y = "Mean Length of Utterance", color = "Utility Condition")+ 
  scale_color_bay() + theme_jmank() +theme(legend.position = "bottom")

# MLU including trials with 0 "length" utterances
expand_utt <- d.raw_chat.final %>%
  select(gameId, trialNum, playerId) %>% 
  distinct() %>%
  full_join(utterances)

@@

#points increase (slightly) over time
d.by_round_metrics %>%
  filter(languageCondition == "Lang") %>%
  ggplot(aes(x = trialNum, y = pointsAdjusted, color = condition)) + 
  geom_point(alpha = .5) + 
  geom_smooth(method=lm)+
  scale_color_bay() + theme_jmank()

```
Point x Trial correlation
```{r}
library(psych)
corel <- d.by_round_metrics %>%
  filter(languageCondition == "Lang") %>% ungroup() %>%
  select(trialNum, pointsAdjusted, numWords)
corr.test(corel$trialNum, corel$pointsAdjusted)
```

@@

```{r}
library(patchwork)
utt_mlu
lang_by_points
rl_comparison
lang_reduction
